---
title: "Untitled"
author: "Chunzi Wang"
date: "April 16 2018"
output: html_document
---
-----

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
library(scales)
library(tidyr)
library(DT)
library(ggthemes)
```

Load in Uber pickup data from 2014 April - 2014 September.

```{r}
apr <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/uber-raw-data-apr14.csv")
may <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/uber-raw-data-may14.csv")
jun <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/uber-raw-data-jun14.csv")
jul <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/uber-raw-data-jul14.csv")
aug <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/uber-raw-data-aug14.csv")
sep <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/uber-raw-data-sep14.csv")
```

```{r}
dim(apr)
str(apr)
summary(apr)
```

Comibe these 6 months of data together.

```{r}
uber_2014 <- rbind(apr,may,jun,jul,aug,sep)
```

Change Date.Time column from factor type to datetime object.

```{r}
uber_2014$Date.Time <- as.POSIXct(uber_2014$Date.Time,format="%m/%d/%Y %H:%M:%S")
```

```{r}
summary(uber_2014)
```

Create new columns of year, month, day, day of week, hour, minute, and second.

```{r}
uber_2014$day <- factor(day(uber_2014$Date.Time))
uber_2014$month <- factor(month(uber_2014$Date.Time, label = TRUE))
uber_2014$year <- factor(year(uber_2014$Date.Time))
uber_2014$dayofweek <- factor(wday(uber_2014$Date.Time, label = TRUE))
```

```{r}
summary(uber_2014)
```

```{r}
uber_2014$Time <- format(as.POSIXct(uber_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
uber_2014$hour <- factor(hour(hms(uber_2014$Time)))
uber_2014$minute <- factor(minute(hms(uber_2014$Time)))
uber_2014$second <- factor(second(hms(uber_2014$Time)))
```

```{r}
summary(uber_2014)
```

## Trips every hour

```{r}
uber_2014 %>%
  count(hour) %>%
  ggplot(aes(x=hour,y=n,fill="pink")) +
  geom_col(show.legend = FALSE) +
  ggtitle("Trips Every Hour") +
  scale_y_continuous(labels=comma) +
  ylab("Total Number")
```

Findings:

+ Two peaks: morning rush hour and evening rush hour.
+ There're increasing number of trips since afternoon, and reach summit at 5pm.
+ More trips taking place in the evening than morning.

```{r}
by_hour <- uber_2014 %>%
  group_by(hour) %>%
  summarize(Total = n())

# display an interactive table
datatable(by_hour)
```

## Trips every hour by month

```{r}
colors <- c("#E69F00","#56B4E9","#660066","#FF6600","#0072B2","#006633","#9999CC")
```

```{r}
uber_2014 %>%
  count(hour,month) %>%
  ggplot(aes(x=hour,y=n,fill=month)) +
  geom_col() +
  ggtitle("Trips Every Hour by Month") +
  scale_y_continuous(labels=comma) +
  ylab("Total Number") +
  scale_fill_manual(values=colors)
```

Findings:

+ April, May, and June have similar numbers for every hour.
+ JUly and August have similar numbers for every hour. 
+ September seems to have the largest share among the six months. July and August have a larger share than April, May, and June.

## Trips distribution on day of weeks

```{r}
uber_2014 %>%
  count(dayofweek) %>%
  ggplot(aes(x=dayofweek,y=n)) +
  geom_col(fill="orange") +
  scale_y_continuous(labels=comma) +
  ylab("Total Number") +
  geom_text(aes(label=n),vjust=1.6,color="white",position=position_dodge(0.9),size=4)
```

Thursday and Friday have the highest number of completed trips. Sunday has the lowest. Weekends don't necessarily beat weekdays here.

## Trips by month day

```{r}
uber_2014 %>%
  count(day) %>%
  ggplot(aes(x=day,y=n,fill=n)) +
  geom_col()
```

## Trips by month

```{r}
uber_2014 %>%
  count(month) %>%
  ggplot(aes(x=month,y=n,fill=month)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels=comma) +
  ylab("Total Number") +
  geom_text(aes(label=n),vjust=1.6,color="white",position=position_dodge(0.9),size=4)
```

There's an increasing number of trips by month. (There's gradual increase in the number of trips from April to September.) September is the best month. 

## Trips by day and month

```{r}
uber_2014 %>%
  count(month,dayofweek) %>%
  ggplot(aes(month, n, fill = dayofweek)) + 
  geom_bar( stat = "identity", position = "dodge") +
  ggtitle("Trips by Day and Month") +
  ylab("Total Number") +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values=colors)
```

## Heatmap by hour and day

```{r}
uber_2014 %>%
  count(day,hour) %>%
  ggplot(aes(x=day,y=hour,fill=n)) +
  geom_tile(color="white") +
  ggtitle("Heat Map by Hour and Day")
```

It's obvious that there're two major peak hours: in the morning between 7-8am and in the afternoon and evening from 4-9pm.

## Trips per base

```{r}
uber_2014 %>%
  count(Base) %>%
  ggplot(aes(x="",y=n,fill=Base)) +
  geom_bar(width=1,stat="identity") +
  coord_polar("y",start=0)
```

Base B02598, B02617, B02682 have better business performance than B02512 abd B02764.

## Heatmap by logitude and latitude

```{r}
min_lat <- min(uber_2014$Lat)
max_lat <- max(uber_2014$Lat)
min_lon <- min(uber_2014$Lon)
max_lon <- max(uber_2014$Lon)
```

```{r}
min_lat
max_lat
min_lon
max_lon
```

```{r}
uber_2014 %>%
  ggplot(aes(x=Lon,y=Lat)) +
  geom_point(size=0.6,color="red",alpha=0.5) +
  scale_x_continuous(limits=c(min_lon,max_lon)) +
  scale_y_continuous(limits=c(min_lat,max_lat)) +
  theme_map() +
  ggtitle("NYC Map based on uber rides during 2014 (Apr-Sep)")
```

The realm is larger than the city, we'll shrink the range when plotting the heatmap based on bases.

## Heatmap by base

```{r}
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004
```

```{r}
uber_2014 %>%
  ggplot(aes(x=Lon, y=Lat, color = Base)) +
  geom_point(size=1,alpha=0.5) +
  scale_x_continuous(limits=c(min_long, max_long)) +
  scale_y_continuous(limits=c(min_lat, max_lat)) +
  theme_map() +
  ggtitle("NYC map based on uber rides during 2014 (Apr-Sep) by base")
```

## Line chart to show the change in trips

```{r}
uber_2014$Date <- format(as.POSIXct(uber_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%m/%d/%Y")
```

```{r}
date <- uber_2014 %>%
  count(Date)

head(date)
str(date)
```

```{r}
date$Date <- as.Date(date$Date,"%m/%d/%Y")
colnames(date) <- c("Date","Trip")
```

```{r}
date %>%
  ggplot(aes(x=Date,y=Trip)) +
  geom_line() +
  geom_point()
```

There's a strong seasonality pattern between trips and date.









Another dataset:

```{r}
uber_jan_feb <- read.csv("C:/Users/Adimn/Desktop/uber-pickups-in-new-york-city/Uber-Jan-Feb-FOIL.csv")
```











